Automated neuroimaging data management with the Sea filesystem

### 2021-10-08 paper/paper_neuro.tex
part of sea intro

In order for a file system to be usable by the average researcher on an HPC system, they must be loadable without administrative
privileges. Furthermore, as the applications are typically made to interact with POSIX-based file systems, the new file system must
also be compliant to the format. One method to ensure that these conditions are met is by using the \texttt{LD\_PRELOAD} trick. This trick
enables the interception of library calls by redefining them.

In this paper, we present Sea, a file system designed to automate data management of neuroimaging pipelines running on HPC systems.

\section{Related Work}

Put in intro.
summary of computing landscape for neuroimaging.
HCP, CBRAIN, nipype, joblib. sea can be used in all these contexts.

\section{Materials and Methods}
\subsection{Datasets}
Functional magnetic resonance imaging datasets can vary greatly in total number
of images and number of volumes within each image. To adequately capture the 
diversity of datasets and when Sea could be pertinent, we selected three datasets of varying
sizes: 1) OpenNeuro's ds001545 dataset~\cite{ds001545}, 2) the PREVENT-AD  dataset~\cite{preventad}, and
3) the HCP dataset~\cite{hcp}. The ds001545 dataset is a total of 45.94~$GB$ (1778 files) and consists of data collected
from 30 subjects in a single session. The PREVENT-AD dataset is a 255.0~$GB$ dataset (53061 files) consisting of data collected
from 308 subjects. The HCP dataset is the largest of the three datasets at 85.4~$TB$ (x files) consisting of data collected from
1113 subjects.

@@

%# files written and output size, duration and memory consumption.
\subsection{fMRI Preprocessing Pipelines}

@@

For AFNI preprocessing, the

>>>>

In order for a file system to be usable by the average researcher on an HPC system, they must be loadable without administrative
privileges. Furthermore, as the applications are typically made to interact with POSIX-based file systems, the new file system must
also be compliant to the format. One method to ensure that these conditions are met is by using the \texttt{LD\_PRELOAD} trick. This trick is
used to intercept select library calls and redefine their behaviour. It has been used in many projects and to create lightweight versions
of filesystems~\cite{xtreemfs}.

In this paper, we present Sea, a file system designed to automate data management of neuroimaging pipelines running on HPC systems.
The goal of Sea is to complement existing neuroimaging applications such that both efficient compute and data management can be achieved
executed on HPC without significant user intervention. As a secondary goal, Sea aims to ensure the fair sharing of cluster usage
by alleviating the impacts of heavy writers on the shared file systems as a whole.

% \section{Related Work}

% Put in intro.
% summary of computing landscape for neuroimaging.
% HCP, CBRAIN, nipype, joblib. sea can be used in all these contexts.

\section{Materials and Methods}

\subsection{The Sea filesystem}

The Sea file system is an on-demand file system that leverages the \texttt{LD\_PRELOAD} trick to intercept POSIX file system (more specifically, glibc)
calls on Linux systems. This enables Sea to redirect write calls aimed for a slower storage device to a faster device whenever
possible. Similarly, when intercepting read calls, Sea can choose to read from a faster device if a copy is available on that device.
Sea decides which storage location it can write to based on the detailed provided in an initialization file called \texttt{sea.ini}. 
This file informs Sea of which locations it can use to read and write to, and well as their order of priority. An example of the initialization
file can be seen in ~\ref{seaini}.

As neuroimaging pipeline results are typically required post-processing and HPC compute-local resources are only accessible during the
reserved duration, Sea provides functionality to flush and evict data to persistent shared storage. This is accomplished via a separate
thread (known as the ``flusher'') that moves data from the caches to long-term storage. A separate thread is used, in this case, to avoid
interrupting on going processing with data management operations.
Users must inform Sea of files that need to be
persisted to storage within a file called \texttt{.sea\_flushlist}, and temporary files which can be removed from cache within a file
called the \texttt{.sea\_evictlist}. Both these files can be populated using regular expressions to denote the paths to flush and evict.
If a file occurs in both the \texttt{.sea\_flushlist} and \texttt{.sea\_evictlist}, Sea will interpret this as a \texttt{move} operation
and simply move the file from local to persistent storage rather than copying it. Files that will be reused by the application should only
be flushed rather than flushed and evicted, as files can benefit from speedups related to reading from fast rather than slow storage.
Sea currently also provides a rudimentary prefetch thread that can move files located within the Sea filesystem to the fastest available cache.
To use Sea's prefetch capabilities, a file called \texttt{.sea\_prefetchlist} needs to be populated using regular expressions like the
flushing and eviction files.

To interact with the Sea file system, a mountpoint is used. The mountpoint is an empty directory that behaves as a view to
all the files and directories stored within Sea. In order to keep track of the locations of the files within the mountpoint, Sea
mirrors the file structure of each storage location across all storage locations. That means, it is generally advisable to provide
empty storage locations for Sea to write to as the mirroring of large directories can take some time. When prefetching, it is
best to provide Sea with a path that only points to that data the requires prefetching as mirroring the paths of full datasets may 
extend processing time.

Sea can easily be launched directly using the available containers on the GitHub registry, or can be compiled via source using Make.
It requires a version of glibc greater than X. Once compiled, Sea can be executed using the `sea-app.sh' binary.

\subsection{Performance analysis of Sea for neuroimaging}

To determine the performance gain Sea brings to neuroimaging analyses, we must evaluate the value of Sea on a variety of neuroimaging
applications. For our analysis, we selected different fMRI preprocessing applications as fMRI processing has some of the most
well-established tools for neuroimaging and some of the largest datasets. Of course, different modalities an tools may result in
vastly different data access patterns and compute times, we use fMRI prepreprocessing as our baseline for the benefits that can be
obtained with Sea.

\subsubsection{Datasets}
Functional magnetic resonance imaging datasets can vary greatly in total number
of images and number of volumes within each image. To adequately capture the 
diversity of datasets and when Sea could be pertinent, we selected three datasets of varying
sizes: 1) OpenNeuro's ds001545 dataset~\cite{ds001545}, 2) the PREVENT-AD  dataset~\cite{preventad}, and
3) the HCP dataset~\cite{hcp}. The ds001545 dataset is a total of 45.94~$GB$ (1778 files) and consists of data collected
from 30 subjects in a single session. The PREVENT-AD dataset is a 255.0~$GB$ dataset (53061 files) consisting of data collected
from 308 subjects. The HCP dataset is the largest of the three datasets at 85.4~$TB$ (x files) consisting of data collected from
1113 subjects.

@@

%# files written and output size, duration and memory consumption.
\subsubsection{fMRI Preprocessing Pipelines}

@@

For AFNI preprocessing, the pipeline was configured to perform slice timing, alignment to Talaraich space, registration, smoothing
and brain masking. While AFNI does have some parallelized components that can be controlled via environment variables, we chose to let 
the pipeline use as much parallelism as required for its execution.

We set fMRIPrep to run with \texttt{--fs-no-reconall} flag set and set the \texttt{--bids-database-dir} to a tmpfs location. 
Furthermore, the seeds were fixed to ensure reproducible results. Unlike the other tools where only a single fMRI image was preprocessed
per process, it was not possible to do so with fMRIPrep. Thus, fMRIPrep preprocessed entire subject directories at a time
